### 2017-05-12 paper/parrows.tex
added figures for the Par Monad chapter.

\subsection{Multicore Haskell}
The Multicore Haskell implementation of this class is implemented in a straightforward manner by using listApp from appendix \ref{utilfns} combined with the \code{withStrategy :: Strategy a -> a -> a} and \code{pseq :: a -> b -> b} combinators from Multicore Haskell, where \code{withStrategy} is the same as \code{using :: a -> Strategy a -> a} but with flipped parameters.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b () where
    	parEvalN _ fs =
       		listApp fs >>>
        	arr (withStrategy (parList rdeepseq)) &&& arr id >>>
        	arr (uncurry pseq)
\end{lstlisting}% $ %% formatting
\caption{Fully evaluating ArrowParallel instance for the Multicore backend}
\label{fig:ArrowParallelMulticoreRdeepseq}
\end{figure}
For most cases a fully evaluating version like in Fig.~\ref{fig:ArrowParallelMulticoreRdeepseq} would probably suffice, but as the Multicore Haskell interface allows the user to specify the level of evaluation to be done via the \code{Strategy} interface, we want to the user not to lose this ability because of using our API. We therefore introduce the \code{Conf a} data-type that simply wraps a \code{Strategy a} (Fig.~\ref{fig:confa}). We can't directly use the \code{Strategy a} type here as GHC (at least in the versions used for development in this paper) does not allow type synonyms in type class instances.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
data Conf a = Conf (Strategy a)
\end{lstlisting}
\caption{Definition of Conf a}
\label{fig:confa}
\end{figure}
To get our configurable \code{ArrowParallel} instance, we simply unwrap the strategy and pass it to \code{parList} as in the fully evaluating version (Fig.~\ref{fig:ArrowParallelMulticoreConfigurable}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b (Conf b) where
    	parEvalN (Conf strat) fs =
        	listApp fs >>>
        	arr (withStrategy (parList strat)) &&& arr id >>>
        	arr (uncurry pseq)
\end{lstlisting}
\caption{Configurable ArrowParallel instance for the Multicore backend}
\label{fig:ArrowParallelMulticoreConfigurable}
\end{figure}
\subsection{ParMonad}
The ParMonad implementation makes use of Haskells laziness and ParMonad's \code{spawnP :: NFData a =>\ \ a -> Par (IVar a)} function. The latter forks away the computation of a value and returns an \code{IVar} containing the result in the \code{Par} monad.

We therefore apply each function to its corresponding input value with \code{app} and then fork the computation away with \code{arr spawnP} inside a \code{zipWithArr} call. This yields a list \code{[Par (IVar b)]}, which we then convert into \code{Par [IVar b]} with \code{arr sequenceA}. In order to wait for the computation to finish, we map over the \code{IVar}s inside the ParMonad with \code{arr (>>= mapM get)}. The result of this operation is a \code{Par [b]} from which we can finally remove the monad again by running \code{arr runPar} to get our output of \code{[b]}.
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b conf where
		parEvalN _ fs = 
			(arr $ \as -> (fs, as)) >>>
			zipWithArr (app >>> arr spawnP) >>>
			arr sequenceA >>>
			arr (>>= mapM get) >>>
			arr runPar
\end{lstlisting} %$ %% formatting

\subsection{Eden}
For the Multicore and ParMonad implementation we could use general instances of \code{ArrowParallel} that just require the \code{ArrowApply} and \code{ArrowChoice} typeclasses. With Eden this is not the case as we can only spawn a list of functions and we cannot extract simple functions out of arrows. While we could still manage to have only one class in the module by introducing a typeclass like
\begin{lstlisting}[frame=htrbl]
class (Arrow arr) => ArrowUnwrap arr where
	arr a b -> (a -> b)
\end{lstlisting} %
we don't do it here, for aesthetic resons. For now, we just implement \code{ArrowParallel} for normal functions
\begin{lstlisting}[frame=htrbl]
instance (Trans a, Trans b) => ArrowParallel (->) a b conf where
parEvalN _ fs as = spawnF fs as
\end{lstlisting}
and the Kleisli type.
\begin{lstlisting}[frame=htrbl]
instance (Monad m, Trans a, Trans b, Trans (m b)) =>
	ArrowParallel (Kleisli m) a b conf where
parEvalN conf fs =
	(arr $ parEvalN conf (map (\(Kleisli f) -> f) fs)) >>>
	(Kleisli $ sequence)
\end{lstlisting}

>>>>

\subsection{Multicore Haskell}
The Multicore Haskell implementation of this class is implemented in a straightforward manner by using listApp from appendix \ref{utilfns} combined with the \code{withStrategy :: Strategy a -> a -> a} and \code{pseq :: a -> b -> b} combinators from Multicore Haskell, where \code{withStrategy} is the same as \code{using :: a -> Strategy a -> a} but with flipped parameters.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b () where
    	parEvalN _ fs =
       		listApp fs >>>
        	arr (withStrategy (parList rdeepseq)) &&& arr id >>>
        	arr (uncurry pseq)
\end{lstlisting}% $ %% formatting
\caption{Fully evaluating ArrowParallel instance for the Multicore backend}
\label{fig:ArrowParallelMulticoreRdeepseq}
\end{figure}
For most cases a fully evaluating version like in Fig.~\ref{fig:ArrowParallelMulticoreRdeepseq} would probably suffice, but as the Multicore Haskell interface allows the user to specify the level of evaluation to be done via the \code{Strategy} interface, we want to the user not to lose this ability because of using our API. We therefore introduce the \code{Conf a} data-type that simply wraps a \code{Strategy a} (Fig.~\ref{fig:confa}). We can't directly use the \code{Strategy a} type here as GHC (at least in the versions used for development in this paper) does not allow type synonyms in type class instances.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
data Conf a = Conf (Strategy a)
\end{lstlisting}
\caption{Definition of Conf a}
\label{fig:confa}
\end{figure}
To get our configurable \code{ArrowParallel} instance, we simply unwrap the strategy and pass it to \code{parList} as in the fully evaluating version (Fig.~\ref{fig:ArrowParallelMulticoreConfigurable}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b (Conf b) where
    	parEvalN (Conf strat) fs =
        	listApp fs >>>
        	arr (withStrategy (parList strat)) &&& arr id >>>
        	arr (uncurry pseq)
\end{lstlisting}
\caption{Configurable ArrowParallel instance for the Multicore backend}
\label{fig:ArrowParallelMulticoreConfigurable}
\end{figure}
\subsection{Par Monad}
The ParMonad implementation (Fig.~\ref{fig:ArrowParallelParMonad}) makes use of Haskells laziness and ParMonad's \code{spawnP :: NFData a =>\ \ a -> Par (IVar a)} function. The latter forks away the computation of a value and returns an \code{IVar} containing the result in the \code{Par} monad.

We therefore apply each function to its corresponding input value with \code{app} and then fork the computation away with \code{arr spawnP} inside a \code{zipWithArr} call. This yields a list \code{[Par (IVar b)]}, which we then convert into \code{Par [IVar b]} with \code{arr sequenceA}. In order to wait for the computation to finish, we map over the \code{IVar}s inside the ParMonad with \code{arr (>>= mapM get)}. The result of this operation is a \code{Par [b]} from which we can finally remove the monad again by running \code{arr runPar} to get our output of \code{[b]}.
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b conf where
		parEvalN _ fs = 
			(arr $ \as -> (fs, as)) >>>
			zipWithArr (app >>> arr spawnP) >>>
			arr sequenceA >>>
			arr (>>= mapM get) >>>
			arr runPar
\end{lstlisting} %$ %% formatting
\caption{ArrowParallel instance for the Par Monad backend}
\label{fig:ArrowParallelParMonad}
\end{figure}

\subsection{Eden}
For the Multicore and ParMonad implementation we could use general instances of \code{ArrowParallel} that just require the \code{ArrowApply} and \code{ArrowChoice} typeclasses. With Eden this is not the case as we can only spawn a list of functions and we cannot extract simple functions out of arrows. While we could still manage to have only one class in the module by introducing a typeclass \code{ArrowUnwrap} (Fig.~\ref{fig:ArrowUnwrap}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
class (Arrow arr) => ArrowUnwrap arr where
	arr a b -> (a -> b)
\end{lstlisting}
\caption{possible ArrowUnwrap typeclass}
\label{fig:ArrowUnwrap}
\end{figure}
We don't do it here for aesthetic resons, though. For now, we just implement \code{ArrowParallel} for normal functions (Fig.~\ref{fig:ArrowParallelEdenFns})
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (Trans a, Trans b) => ArrowParallel (->) a b conf where
parEvalN _ fs as = spawnF fs as
\end{lstlisting}
\caption{ArrowParallel instance for functions in the Eden backend}
\label{fig:ArrowParallelEdenFns}
\end{figure}
and the Kleisli type (Fig.~\ref{fig:ArrowParallelEdenKleisli}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
instance (Monad m, Trans a, Trans b, Trans (m b)) =>
	ArrowParallel (Kleisli m) a b conf where
parEvalN conf fs =
	(arr $ parEvalN conf (map (\(Kleisli f) -> f) fs)) >>>
	(Kleisli $ sequence)
\end{lstlisting}
\caption{ArrowParallel instance for the Kleisli type in the Eden backend}
\label{fig:ArrowParallelEdenKleisli}
\end{figure}
